Attribute fusion in a latent process model for time series of graphs

### 2011-04-29 ssp2011.tex
Add discussion of scan statistic

@@

  \mathbf{1}\{ \phi(uv,t) = k \} \\

@@

  \sum_{u \in N(v)}{\mathbf{1}\{\phi(uv,t) = k\}} \\

@@

  \lambda_k \!\!\sum_{u,w \in N(v)}\!\!\! \mathbf{1}\{\phi(uw,t) = k\} 

@@

@@

@@

@@

@@

@@

    d_{\lambda}(v;t) = \sum_{k = 1}^{K} \lambda_k \sum_{u \in N(v)}
    \mathbf{1}\{ \phi(uv,t) = k \}

@@

  By applying the multivariate normal approximation to the
  multinomial, 

@@

\bibliography{ssp2011}
\end{document}

>>>>

\usepackage{cite}

@@

  \mathbb{I}\{ \phi(uv,t) = k \} \\

@@

  \sum_{u \in N(v)}{\mathbb{I}\{\phi(uv,t) = k\}} \\

@@

  \lambda_k \!\!\sum_{u,w \in N(v)}\!\!\! \mathbb{I}\{\phi(uw,t) = k\} 

@@

    \label{eq:15}

@@

    \label{eq:14}

@@

  \label{eq:13}

@@

\subsection{Power estimates for $\Psi_{\lambda}(t)$}
The limiting distribution for the scan statistics for unattributed
random graphs was considered in
\cite{rukhin:_limit_distr_graph_scan_statis}. This subsection is
concerned with the extension of the results in
\cite{rukhin:_limit_distr_graph_scan_statis} to attributed random
graphs. We state here some the relevant results. Sketches of the
proofs of these results will be provided in
\S~\ref{sec:proofs-some-stated}. \\ \\
Let $\psi_{\lambda}(v;t)$ be the (fused) locality statistics for
vertex $v$ at time $t$, i.e.,
  \label{eq:11}
  \psi_{\lambda}(v;t) = \sum_{k=1}^{K} \lambda_k \sum_{uw \in N(v)}
  \mathbb{I}\{\phi(uw,t) = k\}.
We consider first the case where $t < t^{*}$. Then, conditional on
$|N(v)| = l$, we have
  \psi_{\lambda}(v;t) = \sum_{k=1}^{K}{\lambda_k (y_k + z_k)}
where $(y_1, \dots, y_K)$ and $(z_1, \dots, z_K)$ are distributed as
  (y_1,y_2,\dots,y_K) \sim \textrm{multinomial}\Bigl(
  \tbinom{l}{2}, \pi_{00}\Bigr) \\ (z_1, z_2, \dots, z_K) \sim
  \textrm{multinomial}\Bigl(l, \frac{\pi_{00}}{\langle \pi_{00},
    1\rangle}\Bigr). 
Therefore, for sufficiently large $l$, by the multivariate central
limit theorem, we have
  \Bigl(\psi_\lambda(v;t) \, \lvert \, |N(v)| = l \Bigr) \sim
  \mathcal{N}\Bigl(\tbinom{l}{2} \langle \lambda, \pi_{00} \rangle,
  \tbinom{l}{2} \langle \lambda, \eta_{00} \lambda \rangle \Bigr)
Let $\lambda^{(2)}$ be the vector whose entries are element-wise
square of $\lambda$. Then it can be checked that 
 \langle \lambda,
\eta_{00} \lambda \rangle = \langle \lambda^{(2)}, \pi_{00} \rangle -
  (\langle \lambda, \pi_{00} \rangle)^2
Suppose that $\eta_{00}$ is positive definite. We then have the
following result
  \label{prop:1}
  Let $C = \tfrac{\langle \lambda^{(2)}, \pi_{00} \rangle}{\langle
    \lambda, \pi_{00}\rangle}$ and $p = \tfrac{(\langle \lambda, \pi_{00}
    \rangle)^{2}}{\langle x^{(2)} \pi_{00} \rangle}$. Then $p < 1$,
  and for sufficiently large $l$ we have
  \label{eq:17}
 \Bigl(\psi_\lambda(v;t) \,\lvert\,\, |N(v)| = l \Bigr) \sim C *  
 \mathrm{Bin}\Bigl(\tbinom{l}{2}, p\Bigr)
Proposition~\ref{prop:1} states that the locality
statistics for the attributed random graphs model under consideration
can be approximated by the locality statistics for an
Erd\"{o}s-Renyi graph with edge probability $p$. By a result in
\cite{rukhin:_limit_distr_graph_scan_statis}, we have
  \label{thm:2}
  Let $\Psi_{\lambda}(t)$ be the (fused) scan statistic. Let
  $N_{0}$, $a_n$, and $b_n$ be defined by
   \mu_0 &= (n-1) \langle \pi_{00}, 1 \rangle \\
 \sigma_0 &= \sqrt{(n-1) (\langle \pi_{00}, 1 \rangle - \langle
   \pi_{00}, 1 \rangle^2)} \\ 
 z_n &= \sqrt{2
      \log{n}}\Bigl(1 - \frac{\log{\log{n}} + \log{4 \pi}}{4
      \log{n}}\Bigr) \\
    N_{0} &= \mu_0 + z_n \sigma_0 \\
   a_n &= N_{0} \frac{\langle \lambda, \pi_{00} \rangle}{\langle 1,
     \pi_{00} \rangle} + p\tbinom{N_{0}}{2} \\ 
   b_n &= (1 - \tfrac{p}{2} + p N_{0}) \frac{\sigma_0}{\sqrt{2 \log
     n}}
Then we have
  \Psi_{\lambda}(t) \rightsquigarrow \mathcal{G}(C a_n, C b_n)
We now move to the case where $t = t^{*}$. Let $d_{\zeta}(v;t)$ be
the number of neighbors of $v$ that lies in $[m]$ (the egg) and
$d_{\xi}(v;t)$ be the number of neighbors of $v$ that lies in $[n]
\setminus [m]$ (the kidney). Conditional on $d_{\zeta}(v;t^{*}) =
l_{\zeta}$, $d_{\xi}(v;t^{*}) = l_{\xi}$, we have
  \psi_{\lambda}(v;t^{*}) = \sum_{k=1}^{K} \lambda_k ( y^{(\zeta)}_k + y^{(\xi)}_k +
  y^{(\omega)}_k + z^{(\zeta)}_k + z^{(\xi)}_k)
where $(y^{(\zeta)}_1, \dots, y^{(\zeta)}_m)$, $(y^{(\xi)}_1,\dots,
 y^{(\xi)}_m)$, $(y^{(\omega)}_1, \dots, y^{(\omega)}_m)$ are
 distributed as
(y^{(\zeta)}_1,\dots,y^{(\zeta)}_m) \sim \textrm{multinomial}\Bigl(
\tbinom{l_\zeta}{2}, \pi_{11}\Bigr) \\ 
(y^{(\xi)}_1,\dots,y^{(\xi)}_m) \sim \textrm{multinomial}\Bigl(
\tbinom{l_\xi}{2}, \pi_{00}\Bigr) \\
(y^{(\omega)}_1,\dots,y^{(\omega)}_m) \sim \textrm{multinomial}\Bigl(
l_\zeta l_\xi, \pi_{10}\Bigr)
and $(z^{(\zeta)}_1, \dots, z^{(\zeta)}_m)$, $(z^{(\xi)}_1, \dots,
z^{(\xi)}_m)$ are distributed as
(z^{(\zeta)}_1, \dots, z^{(\zeta)}_m) \sim
\begin{cases}
\textrm{multinomial}\Bigl(l_\zeta, \frac{\pi_{11}}{\langle \pi_{11},
  1\rangle}\Bigr) & \text{if $v \in [m]$} \\
\textrm{multinomial}\Bigl(l_\zeta, \frac{\pi_{10}}{\langle \pi_{10}, 1
  \rangle} \Bigr) & \text{if $v \not\in [m]$}
\end{cases} \\
(z^{(\xi)}_1, \dots, z^{(\xi)}_m) \sim
\begin{cases}
\textrm{multinomial}\Bigl(l_\xi, \frac{\pi_{10}}{\langle \pi_{10},
  1\rangle}\Bigr) & \text{if $v \in [m]$} \\
\textrm{multinomial}\Bigl(l_\zeta, \frac{\pi_{00}}{\langle \pi_{00}, 1
  \rangle} \Bigr) & \text{if $v \notin [m]$}
Thus, for sufficiently large $l_{\zeta}$ and $l_{\xi}$, we have
 \psi_{\lambda}(v;t^{*}) & \sim  \mathcal{N}\Bigl(\tbinom{l_\zeta}{2}
 \langle \lambda, \pi_{11} \rangle,
  \tbinom{l_\zeta}{2} \langle x, \eta_{11} \lambda \rangle \Bigr) \\ &+ 
  \mathcal{N}\Bigl(\tbinom{l_\xi}{2} \langle \lambda, \pi_{00} \rangle,
  \tbinom{l_\xi}{2} \langle \lambda, \eta_{00} \lambda \rangle \Bigr) \\ &+ 
  \mathcal{N}\Bigl(l_\zeta l_\xi \langle \lambda, \pi_{10} \rangle,
  l_\zeta l_\xi \langle \lambda, \eta_{10} \lambda \rangle \Bigr)
  \label{prop:2}
  Let $C_1, C_2, C_3$ and $p_1$, $p_2$, $p_3$ be 
C_1 = \tfrac{\langle \lambda^{(2)}, \pi_{00} \rangle}{\langle \lambda,
    \pi_{00}\rangle}; \quad p_1 = \tfrac{(\langle \lambda, \pi_{00}
    \rangle)^{2}}{\langle \lambda^{(2)} \pi_{00} \rangle} \\
C_2 = \tfrac{\langle \lambda^{(2)}, \pi_{11} \rangle}{\langle \lambda,
    \pi_{11}\rangle}; \quad p_2 = \tfrac{(\langle \lambda, \pi_{11}
    \rangle)^{2}}{\langle \lambda^{(2)} \pi_{11} \rangle} \\
C_3 = \tfrac{\langle \lambda^{(2)}, \pi_{11} \rangle}{\langle \lambda,
    \pi_{10}\rangle}; \quad p_1 = \tfrac{(\langle \lambda, \pi_{10}
    \rangle)^{2}}{\langle \lambda^{(2)} \pi_{10} \rangle} 
  Then $p_1$, $p_2$, and $p_3$ are all at most $1$. Furthermore, for
  sufficiently large $l_\zeta$ and $l_\xi$ we have
 \psi_{\lambda}(v;t_1) &\sim C_1 *  
 \mathrm{Bin}\Bigl(\tbinom{l_\zeta}{2}, p_1\Bigr) + C_2 *  
 \mathrm{Bin}\Bigl(\tbinom{l_\xi}{2}, p_2\Bigr) \\ &+ C_3 *
 \mathrm{Bin}\Bigl(l_\zeta l_\xi, p_3\Bigr)
Proposition~\ref{prop:2} state that the locality statistics
for the attributed random graphs model under consideration at time $t
= t^{*}$ can be approximated by the locality
statistics for an unattributed kidney and egg random graph model
$\kappa(n,n_1,p_1,p_2,p_3)$. Let us define $B$, $C$, $E$ and $F$ by
  B \sim \mathrm{Bin}(m, \langle 1, \pi_{00} \rangle); \quad C \sim \mathrm{Bin}(n-m-1,
 \langle 1, \pi_{01} \rangle) \\
  E \sim \mathrm{Bin}(m-1, \langle 1, \pi_{11} \rangle); \quad F \sim \mathrm{Bin}(n-m,
  \langle 1, \pi_{01} \rangle)
For any random variable $X$, we let $\mu_X$ and $\sigma_X^{2}$ denote
the mean and variance of $X$. If $X$ and $Y$ are independent random
variables, we let $\mu_{X + Y}$ and $\sigma_{X + Y}^2$ denote the mean
and variance of the convolution $X + Y$. \\
\noindent Let $N_{\kappa}$ be the quantity defined by
  \label{eq:18}
  N_{\kappa} = \mu_{E + F} + z_m \sigma_{E + F}
where $z_m$ is
  z_m = \sqrt{2 \log m}\Bigl(1 - \frac{\log{\log{m}} + \log{4
      \pi}}{\log{m}}\Bigr)
The following result is adapted from \cite{rukhin:_limit_distr_graph_scan_statis}
  \label{thm:3}
  Let $a_{n,m}$ and $b_{n,m}$ be given by
    \label{eq:28}
    a_{n,m} &= N_{\kappa} \frac{\langle x, \pi_{01} \rangle}{\langle 1, \pi_{01} \rangle} \\ &+ C_1 p_1 \binom{N_\kappa}{2} + (C_2 p_2 -
    C_1 p_1)\binom{\mu_E}{2} \\ &+ (C_3 p_3 - C_1 p_1)\mu_E \mu_F 
    \end{split} \\
    b_{n,m} = (1 - \tfrac{C_1 p_1}{2} + C_1 p_1
    N_\kappa)\frac{\sigma_{E + F}}{\sqrt{2 \log{n_1}}}
  If $m = \Omega(\sqrt{n \log n})$, then
    \label{eq:29}
    \Psi_{\lambda}(t^{*}) \rightsquigarrow \mathcal{G}(a_{n,m}, b_{n,m})

@@

\label{sec:proofs-some-stated}

@@

    d_{\lambda}(v;t) = \sum_{k = 1}^{K} \lambda_k \sum_{w \in N(v)}
    \mathbb{I}\{ \phi(vw,t) = k \}

@@

  Let $\Gamma^{(k)}(v;t) = |\{w \colon \mathbb{I}\{\phi(wv,t) =
  k\}|$. The vector $\bm{\Gamma}(v;t) = (\Gamma^{(1)}(v;t), \dots,
  \Gamma^{(K)}(v;t))$ for $t < t^{*}$ is distributed as a multinomial
  with $n-1$ trials and probability vector $\pi_0$. The following
  statements are made in the limit as $n \rightarrow \infty$ for fixed
  $K$. By the multivariate
  central limit theorem, we have
    \bm{\Gamma}(v;t) \sim \mathcal{N}((n-1)\pi_0,
    (n-1)(\mathrm{diag}(\pi_0) - \pi_0 \pi_0^{T}))
  Now, $d_{\lambda}(v;t) = \langle \lambda, \bm{\Gamma}(v;t) \rangle$
  and therefore
    \label{eq:12}
    d_{\lambda}(v;t) \sim \mathcal{N}((n-1) \langle \lambda, \pi_0
    \rangle, (n-1) \langle \lambda, \eta_{00} \lambda \rangle)
  where $\eta_{00} = \mathrm{diag}(\pi_0) - \pi_0
  \pi_0^{T}$. $\Delta_{\lambda}(t)$ for $t < t^{*}$ is thus the
  maximum of a sequence of {\em dependent} Gaussian random
  variables. By an argument analogous to the argument for the maximum
  degree of an Erd\"{o}s-Renyi random graph in \cite[\S
  III.1]{bollobas85:_random_graph} we can show that the dependency
  among the $d_{\lambda}(v,t)$ can be safely
  ignored. $\Delta_{\lambda}(t)$ is then equivalent to the maximum of
  a sequence of {\em independent} $\mathcal{N}((n-1) \langle \lambda,
  \pi_0 \rangle, (n-1) \langle \lambda, \eta_{00} \lambda
  \rangle)$. The resulting distribution is a Gumbel \cite[\S
  2.3]{galambos87:_asymp_theor_extrem_order_statis} and we obtain 
  Eq.~\eqref{eq:14} for $t < t^{*}$. The case when $t = t^{*}$ can be
  derive in a similar manner, see e.g,
  \cite{rukhin:_limit_distr_graph_scan_statis}.

@@

\begin{proof}[Lemma~\ref{lem:2}]
  Let $ X \sim \mathcal{G}(\alpha, \beta)$. We consider the
  normalization $\tfrac{X - \mu}{\sigma}$. We have
    \mathbb{P}[ \frac{X - \mu}{\sigma} \leq z]  &= \mathbb{P}[X \leq z
    \sigma + \mu] 
    = e^{-e^{-(z \sigma + \mu - \alpha)/\beta}} \\
      &= e^{- e^{-(z - (\alpha - \mu)/\sigma)/(\beta/\sigma)}}
  Thus, $\tfrac{X - \mu}{\sigma} \sim \mathcal{G}(\tfrac{\alpha -
    \mu}{\sigma}, \tfrac{\beta}{\sigma})$. Because the sample
  mean and the sample variance are consistent estimators, the claims
  follow after an application of Slutsky's theorem.
\bibliography{ssp2011}
\end{document}
